New lower bounds for Schur and weak Schur numbers

### 2021-06-12 article/article.tex
modifs

This article provides new lower bounds for both Schur and weak Schur numbers. These results were obtained by 
continuing on Rowley's "template"-based approach for Schur and Ramsey numbers. Finding templates allows us to 
get explicit partitions improving lower bounds as well as the growth rate for both Schur numbers and Ramsey 
numbers \(R_n(3)\). We also developed a method to improve lower bounds for weak Schur numbers. 
Furthermore, we try to analyze former works on weak Schur numbers based on the principle that \textit{good} 
partitions into \(n+1\) weakly sum-free subsets start with a \textit{good} partition into \(n\) weakly sum-free subsets. 
We show that exceeding the previous lower bound \(\WS (6) \geqslant 582\) is impossible with such an assumption 
upon imposing certain conditions on the \textit{good} 5-subsets partitions. The new lower bounds include 
\(S(9) \geqslant 17\,803\), \(S(10) \geqslant 60\,948\), \(\WS (9) \geqslant 22\,536\) and 
\(\WS (10) \geqslant 71\,214 \).

>>>>

This article provides new lower bounds for both Schur and weak Schur numbers. These results were obtained by 
continuing on Rowley's "template"-based approach for Schur and Ramsey numbers. We also generalized the concept 
of "template" to weak Schur numbers. Finding new templates allows us to get explicit partitions improving lower bounds 
as well as the growth rate for Schur numbers, weak Schur numbers and Ramsey numbers \(R_n(3)\). The new lower 
bounds include \(S(9) \geqslant 17\,803\), \(S(10) \geqslant 60\,948\), \(\WS (9) \geqslant 22\,536\) and 
\(\WS (10) \geqslant 71\,214 \).

### 2021-06-12 article/introduction.tex
modifs

\qquad Before Rowley's "template"-based approach for Schur and Ramsey numbers \cite{RowleyRamsey}, the 
previous generic construction for Schur numbers was given by Abbott and Hanson \cite{AbbottHanson} in 1972 
with a recursive construction. It used to give the best lower bounds for all sufficiently large numbers. No equivalent 
was known for weak Schur numbers and as a result the best known partitions for large weak Schur numbers 
did not use the weakly sum-free hypothesis.

\par
As for smaller weak Schur numbers, the best lower bounds were obtained by conducting a computer search. Eliahou 
\cite{EliahouBook}, Bouzy \cite{Bouzy2015AnAP} and Rafilipojaona \cite{Rafilipojaona} improved the lower 
bounds with Monte-Carlo methods. This was the main approach during the past decade. This search for weakly 
sum-free partitions relied on the recursive assumption that a good weakly sum-free partition into \(n+1\) subsets 
starts with a good weakly sum-free partition into \(n\) subsets.

\par
In 2020, Rowley introduced the notion of templates for Schur and Ramsey numbers \cite{RowleyRamsey} which 
generalizes Abbott and Hanson's construction and produces new lower bounds (and inequalties) for Schur numbers. 
He also provides two inequalities for weak Schur numbers \cite{RowleyWS} that yield significant improvements 
over previous lower bounds, and besides do utilize the \textit{weakly} sum-free hypothesis.

@@

\begin{table}
  \centering
\begin{tabular}{|*{13}{c|}}
    \hline
    \(n\) & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 & 11 & 12 \\
    \hline
    \multirow{2}{*}{Before Rowley} & 1* & 4* & 13* & 44* & 160* & 536 & 1\,680 & 5\,041 & 15\,124 & 51\,120 & 172\,216 & 575\,664 \\
      & & & & & \cite{Heule2017} & \cite{Fredricksen} & \cite{Fredricksen} & \cite{AbbottHanson} & 
    \cite{AbbottHanson} & \cite{AbbottHanson} & \cite{AbbottHanson} & \cite{AbbottHanson} \\
    \hline
    \begin{tabular}{@{}c@{}}Rowley \\ \cite{RowleyRamsey}\end{tabular} & & & & & & & & 5\,286 & 17\,694 & 
    60\,320 & 201\,696 & 631\,840 \\
    \hline
    \hyperref[Schur]{\textbf{Our results}}  & & & & & & & & & \textbf{17\,803} & \textbf{60\,948} & \textbf{203\,828} & \textbf{638\,548} \\
    \hline
\end{tabular}
\caption{Comparison of lower bounds for Schur numbers}
\label{table:1}
\end{table}
\vspace{2.5ex}

\begin{table}
  \centering
\begin{tabular}{|*{13}{c|}}
    \hline
    \(n\) & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 & 11 & 12 \\
    \hline
    \multirow{2}{*}{Before Rowley} & 2* & 8* & 23* & 66* & 196 & 582 & 1\,740 & 5\,201 & 15\,596 & 51\,520 & 172\,216 & 575\,664 \\
      & & & & & \cite{ELIAHOU2012175} &\cite{EliahouBook} & \cite{Rafilipojaona} & \cite{Rafilipojaona} & 
    \cite{Rafilipojaona} & \cite{AbbottHanson} & \cite{AbbottHanson} & \cite{AbbottHanson} \\
    \hline
    \begin{tabular}{@{}c@{}}Rowley \\ \cite{RowleyWS}\end{tabular} & & & & & & 642 & 2\,146 & 6\,976 & 21\,848 
    & 70\,778 & 241\,282 & 806\,786 \\
    \hline
    \hyperref[WeakSchur]{\textbf{Our results}} & & & & & & & & & \textbf{22\,536} & \textbf{71\,214} & \textbf{243\,794} & \textbf{815\,314} \\
    \hline
\end{tabular}

\vspace{1.5ex}
* denotes the exact value
\caption{Comparison of lower bounds for weak Schur numbers}
\label{table:2}
\end{table}

@@

\qquad Our main contribution is a generalization of the concept of template to weak Schur numbers. 
Our templates provide new lower bounds (and inequalities) for weak Schur numbers. This construction also includes, 
as a special case, a construction similar to Abbott and Hanson's \cite{AbbottHanson}, but this time with 
\textit{weak} Schur numbers.

\par
In the \hyperref[Schur]{section 2}, we explain Rowley's template-based construction in the context of 
Schur numbers. Then, we give new templates, thus providing new lowers bounds and inequalities as well as 
showing that the growth rates for both Schur and Ramsey numbers \(R_n(3)\) exceed 3.28.

\par
Then, in the \hyperref[WeakSchur]{section 3}, we generalize the concept of templates to weak Schur numbers 
and provide new lower bounds for weak Schur numbers.

@@

\begin{definition}
Let \(n \in \mathbb{N}^*\). There exists a greatest integer denoted by \(S(n)\) (\textit{resp. \(\WS (n)\)}) such that
\([\![1,S(n)]\!]\) (resp. \([\![1, \WS (n)]\!]\)) can be partitioned into \(n\) sum-free subsets (resp. weakly sum-free
subsets). \(S(n)\) is called the \textit{\(n\)\textsuperscript{th} Schur number} and \textit{\(\WS (n)\) the
\(n\)\textsuperscript{th} weak Schur number}.
\end{definition}

>>>>

\qquad Up to recently, the most efficient generic construction for Schur numbers was given by Abbott and Hanson 
\cite{AbbottHanson} in 1972 with a recursive construction. It gave the best lower bounds for all sufficiently large 
numbers. No equivalent was known for weak Schur numbers and as a result the best known partitions for large 
weak Schur numbers did not use the weakly sum-free hypothesis.

\par
As for smaller weak Schur numbers, the best lower bounds were obtained by conducting a computer search. Eliahou 
\cite{EliahouBook}, Bouzy \cite{Bouzy2015AnAP} and Rafilipojaona \cite{Rafilipojaona} improved the lower 
bounds with Monte-Carlo methods. This was the main approach during the past decade.

\par
In 2020, Rowley introduced the notion of templates for Schur and Ramsey numbers \cite{RowleyRamsey} which 
generalizes Abbott and Hanson's construction and produces new lower bounds (and inequalties) for Schur numbers. 
He also provides two inequalities for weak Schur numbers \cite{RowleyWS} that yield significant improvements 
over previous lower bounds. Besides, these two inequalities do use the \textit{weakly} sum-free hypothesis.

@@

\begin{table}[H]
  \centering
\begin{tabular}{|*{13}{c|}}
    \hline
    \(n\) & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 & 11 & 12 \\
    \hline
    \multirow{2}{*}{Before Rowley} & 1* & 4* & 13* & 44* & 160* & 536 & 1\,680 & 5\,041 & 15\,124 & 51\,120 & 172\,216 & 575\,664 \\
      & & & & & \cite{Heule2017} & \cite{Fredricksen} & \cite{Fredricksen} & \cite{AbbottHanson} & 
    \cite{AbbottHanson} & \cite{AbbottHanson} & \cite{AbbottHanson} & \cite{AbbottHanson} \\
    \hline
    \begin{tabular}{@{}c@{}}Rowley \\ \cite{RowleyRamsey}\end{tabular} & & & & & & & & 5\,286 & 17\,694 & 
    60\,320 & 201\,696 & 631\,840 \\
    \hline
    \hyperref[Schur]{\textbf{Our results}}  & & & & & & & & & \textbf{17\,803} & \textbf{60\,948} & \textbf{203\,828} & \textbf{638\,548} \\
    \hline
\end{tabular}
\caption{Comparison of lower bounds for Schur numbers}
\label{table:1}
\end{table}
\vspace{2.5ex}

\begin{table}[H]
  \centering
\begin{tabular}{|*{13}{c|}}
    \hline
    \(n\) & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 & 11 & 12 \\
    \hline
    \multirow{2}{*}{Before Rowley} & 2* & 8* & 23* & 66* & 196 & 582 & 1\,740 & 5\,201 & 15\,596 & 51\,520 & 172\,216 & 575\,664 \\
      & & & & & \cite{ELIAHOU2012175} &\cite{EliahouBook} & \cite{Rafilipojaona} & \cite{Rafilipojaona} & 
    \cite{Rafilipojaona} & \cite{AbbottHanson} & \cite{AbbottHanson} & \cite{AbbottHanson} \\
    \hline
    \begin{tabular}{@{}c@{}}Rowley \\ \cite{RowleyWS}\end{tabular} & & & & & & 642 & 2\,146 & 6\,976 & 21\,848 
    & 70\,778 & 241\,282 & 806\,786 \\
    \hline
    \hyperref[WeakSchur]{\textbf{Our results}} & & & & & & & & & \textbf{22\,536} & \textbf{71\,214} & \textbf{243\,794} & \textbf{815\,314} \\
    \hline
\end{tabular}
\caption{Comparison of lower bounds for weak Schur numbers}
\label{table:2}
\end{table}

\begin{center}
* denotes the exact value
\end{center}

@@

\qquad Our main contribution is a generalization of the concept of template to weak Schur numbers. Our templates 
provide new lower bounds (and inequalities) for weak Schur numbers. This construction also includes as a special case 
a construction similar to Abbott and Hanson's \cite{AbbottHanson}, but this time for \textit{weak} Schur numbers.

\par
In \hyperref[Schur]{section 2}, we explain Rowley's template-based construction in the context of 
Schur numbers. Then, we give new templates, thus providing new lowers bounds and inequalities as well as 
showing that the growth rates for both Schur and Ramsey numbers \(R_n(3)\) exceed 3.28.

\par
In \hyperref[WeakSchur]{section 3}, we generalize the concept of templates to weak Schur numbers 
and provide new lower bounds for weak Schur numbers.

@@

\begin{definition}
Let \(n \in \mathbb{N}^*\). There exists a largest integer denoted by \(S(n)\) such that \([\![1, S(n)]\!]\) can be
 partitioned into \(n\) sum-free subsets. \(S(n)\) is called the \textit{\(n\)\textsuperscript{th} Schur number}.
\end{definition}

\begin{definition}
Let \(n \in \mathbb{N}^*\). There exists a largest integer denoted by \(\WS (n)\) such that \([\![1, \WS (n)]\!]\)) 
can be partitioned into \(n\) weakly sum-free subsets. \(\WS(n)\) is called the \(n\)\textsuperscript{th} weak Schur 
number.
\end{definition}

### 2021-06-12 article/schur.tex
modifs

\begin{align}
	S(n + 1) \geqslant  3 S(n)  + 1 \label{S(n+1)}\\
	S(n + 2) \geqslant  9 S(n)  + 4 \label{S(n+2)}\\
	S(n + 3) \geqslant  33 S(n) + 6 \label{S(n+3)}\\
	S(n + 4) \geqslant  111 S(n) + 43 \label{S(n+4)}\\
	S(n + 5) \geqslant  380 S(n) + 148 \label{S(n+5)}\\
	S(n + 6) \geqslant 1\,140 S(n) + 528 \label{S(n+6)}
\end{align}

\hyperref[S(n+1)]{(\ref{S(n+1)})} comes from  Schur's original article\cite{Schur1917}. \hyperref[S(n+2)]{(\ref{S(n+2)})} is due to
Abott \cite{AbbottHanson}
and \hyperref[S(n+3)]{(\ref{S(n+3)})} to Rowley \cite{RowleyRamsey}. \hyperref[S(n+4)]{(\ref{S(n+4)})}, \hyperref[S(n+5)]{(\ref{S(n+5)})} and \hyperref[S(n+6)]{(\ref{S(n+6)})} are our result.

\par
\hyperref[S(n+1)]{(\ref{S(n+1)})}, \hyperref[S(n+2)]{(\ref{S(n+2)})} and \hyperref[S(n+3)]{(\ref{S(n+3)})} are optimal. \hyperref[S(n+4)]{(\ref{S(n+4)})} is optimal among symmetric S-templates whose special color is 
the last in the order of apparition (and with a multiplicative factor less than or equal to 118). \hyperref[S(n+5)]{(\ref{S(n+5)})} is most 
likely not optimal but should not be too far from the optimum. 
Finally, \hyperref[S(n+6)]{(\ref{S(n+6)})} is obtained by combining a S-template with width 380 and one with width three. 
Although we could not find a better S-template with seven colors, \hyperref[S(n+6)]{(\ref{S(n+6)})} is definitely very far from the 
optimal value. One may try to seek better templates by constraining less the search space and by using 
Monte-Carlo methods, as in \cite{Bouzy2015AnAP}. This could be the subject of a future work.

>>>>

\begin{align}
	S(n + 1) \geqslant &~3\,S(n)  + 1 \label{S(n+1)}\\
	S(n + 2) \geqslant &~9\,S(n)  + 4 \label{S(n+2)}\\
	S(n + 3) \geqslant &~33\,S(n) + 6 \label{S(n+3)}\\
	S(n + 4) \geqslant &~111\,S(n) + 43 \label{S(n+4)}\\
	S(n + 5) \geqslant &~380\,S(n) + 148 \label{S(n+5)}\\
	S(n + 6) \geqslant &~1\,140\,S(n) + 528 \label{S(n+6)}
\end{align}

Inequality (\ref{S(n+1)}) comes from  Schur's original article\cite{Schur1917}. Inequality (\ref{S(n+2)}) is due to
Abott and Hanson \cite{AbbottHanson} and inequality (\ref{S(n+3)}) to Rowley \cite{RowleyRamsey}. Inequalities 
(\ref{S(n+4)}), (\ref{S(n+5)}) and (\ref{S(n+6)}) are our results.

\par
Inequalities (\ref{S(n+1)}), (\ref{S(n+2)}) and (\ref{S(n+3)}) are optimal. Inequality (\ref{S(n+4)}) is optimal 
among symmetric S-templates whose special color is the last in the order of apparition (and with a multiplicative 
factor less than or equal to 118). Inequality (\ref{S(n+5)}) is most likely not optimal but should not be too far 
from the optimum. Finally, inequality (\ref{S(n+6)}) is obtained by combining a S-template with width 380 and 
one with width three. Although we could not find a better S-template with seven colors, inequality (\ref{S(n+6)}) 
is definitely very far from the optimal value. One may try to seek better templates by constraining less the search 
space and by using Monte-Carlo methods, as in \cite{Bouzy2015AnAP}. This could be the subject of a future work.
